---
title: "Dashboard setup: acquire Comtrade data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{acquire-comtrade-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Storing the API key

A good way of keeping the API key safe is to store it in the .Renviron file

```{r, eval=FALSE}
# .Renviron file setup
usethis::edit_r_environ()

# Your .Renviron file opens for editing
# Text to add, as an example
# COMTRADE_KEY=<text of your key>

# once set, you can retrieve the key with 
Sys.getenv("COMTRADE_KEY")
```

## Acquiring dashboard data

Code below doesnt include every country in the world, and a lot of them are ex countries which could be weeded out.

```{r, eval=FALSE}
# Process for downloading Comtrade data and extracting FFD trade
# Best run as a job in Rstudio

library(iapdashboardadmin)

key <- Sys.getenv("COMTRADE_KEY")
folder <- here::here("data", "downloads")

yrs <- c(2010:2018)
ctys <- c(660, 533, 112, 58, 84, 204, 60, 64, 68, 535, 92, 96, 854, 108, 132,
          116, 136, 140, 148, 344, 446, 174, 178, 184, 384, 192, 531, 200, 408, 180,
          262, 212, 214, 588, 218, 222, 226, 232, 97, 234, 238, 886, 278, 866, 720, 230,
          280, 582, 590, 592, 868, 717, 736, 835, 810, 890, 836, 254, 258, 583, 266,
          270, 268, 288, 292, 304, 308, 312, 320, 324, 624, 328, 332, 336, 340, 356,
          368, 400, 398, 296, 414, 417, 418, 422, 426, 430, 434, 450, 454, 462, 466,
          470, 584, 474, 478, 480, 175, 496, 500, 508, 104, 580, 516, 524, 530, 532,
          540, 558, 562, 512, 490, 585, 598, 600, 459, 634, 498, 638, 646, 647, 461,
          654, 659, 658, 662, 534, 666, 670, 882, 674, 678, 457, 686, 891, 690, 694,
          711,  90, 706, 728, 144, 275, 729, 740, 748, 757, 760, 762, 807, 626, 768,
          772, 776, 780, 788, 795, 796, 798, 800, 826, 834, 858, 850, 841, 860, 548,
          862, 876, 887, 894, 716)

iapdashboardadmin::get_countries_years(periods = yrs,
                                       reporters = ctys,
                                       token = key,
                                       dest_folder = folder,
                                       unzip = FALSE)

ffd_trade <- iapdashboardadmin::merge_ffd(here::here("data", "downloads"))

saveRDS(ffd_trade, here::here("data", "db", paste0(Sys.Date(), "_ffd_trade.rds")))

```

## Data availability summary

Its useful to build a dataset containing a summary of the data that has been acquired

```{r, eval=FALSE}
# Produce a dataframe containing availability of data in the ffd_trade data

library(tidyverse)

reporters <- iapdashboardadmin::reporters
ffd_trade <- readRDS(here::here("data", "db", "ffd_trade.rds"))


data <- ffd_trade %>% 
                select(reporter_code, reporter, year) %>% 
                group_by(reporter_code, reporter, year) %>% 
                summarise(count = TRUE) %>% 
                pivot_wider(names_from = year, values_from = year) %>% 
                select(-count)

availability <- reporters %>% 
      left_join(data, by = c("id" = "reporter_code")) %>% 
      select(-reporter) %>% 
      rename(reporter = text,
             reporter_code = id)

```

## Setting up FFD indicators for the dashboard

We dont want the dashboard to have to work with the large dataset we have extracted from Comtrade dynamically.
Instead we set up the indicators we want and save them in a much smaller dataset.

```{r, eval=FALSE}
library(tidyverse)
library(broom)

ffd_trade <- readRDS(here::here("data", "db", "ffd_trade.rds"))


# Current setup ----------------------------------------------------------------

# Get a list of all the countries in the main dataset
all_data <- ffd_trade %>% 
  group_by(reporter_code, reporter_iso, reporter) %>% 
  summarise(n = n()) %>% 
  select(-n)

# Total food imports and market size rating by year
total_food_imports <- ffd_trade %>% 
  filter(partner_iso == "WLD") %>% 
  group_by(reporter, year) %>% 
  summarise(total_food_imports_value = sum(trade_value_us)) %>% 
  group_by(year) %>% 
  mutate(total_food_imports_rating = ntile(total_food_imports_value, 4)) %>% 
  arrange(year, total_food_imports_value)


# Total UK food imports by year
uk_food_imports <- ffd_trade %>% 
  filter(partner_iso == "GBR") %>% 
  group_by(reporter, year) %>% 
  summarise(uk_food_imports = sum(trade_value_us)) 


uk_market_share <- ffd_trade %>% 
  filter(partner_iso %in% c("WLD", "GBR")) %>% 
  select(year, reporter, partner_iso, commodity, commodity_code, trade_value_us) %>%
  group_by(year, reporter, partner_iso) %>% 
  summarise(trade_value_us = sum(trade_value_us)) %>% 
  pivot_wider(id_cols = c(year, reporter), names_from = partner_iso, values_from = trade_value_us) %>% 
  filter(GBR > 0) %>% 
  mutate(uk_market_share = (GBR/WLD)*100) %>% 
  arrange(reporter, year)

export_diversity <- ffd_trade %>%
  filter(partner_iso == "GBR") %>% 
  group_by(reporter, year, commodity_code) %>% 
  summarise(count = n()) %>% 
  group_by(reporter, year) %>% 
  summarise(uk_export_diversity = sum(count))

dominant_products <- ffd_trade %>%
  filter(partner_iso == "GBR") %>%
  select(year, reporter, commodity, commodity_code, trade_value_us) %>%
  group_by(year, reporter) %>% 
  mutate(uk_percentage = (trade_value_us/sum(trade_value_us)*100)) %>% 
  arrange(year, reporter, desc(uk_percentage)) %>% 
  slice(1)

ffd_indicators %>% 
  filter(reporter == "Argentina") %>% 
  pivot_longer(cols = c(total_food_imports, uk_food_imports, uk_mkt_share_growth)) %>% knitr::kable()

all_data <- all_data %>% 
  left_join(total_food_imports) %>% 
  left_join(uk_food_imports) %>% 
  left_join(uk_market_share) %>% 
  left_join(dominant_products)

# saveRDS(all_data, here::here("data", "db", "ffd_indicators.rds"))


# Models - this isnt working yet -------------------------------------------------------


trade_model <- function(df) {
  lm(uk_market_share ~ year, data = df)
}

mods <- indicators %>% 
  group_by(reporter) %>% 
  nest() %>% 
  mutate(model = map(data, trade_model),
         details = coefficients(model),
         preds = map2(data, model, modelr::add_predictions)) 


preds <- unnest(mods, preds)

preds %>% 
  filter(reporter != "Ireland") %>% 
  ggplot(aes(x = year, colour = reporter)) +
  geom_line(aes(y = pred, group = reporter), alpha = 1 / 3) + 
  geom_line(aes(y = uk_market_share, group = reporter), alpha = 1 / 3) +
  geom_smooth(aes(y = pred), se = FALSE) +
  facet_grid(~ reporter) +
  theme(legend.position = "none")


cncodes <- read_csv(here::here("data", "reference", "CN CODES MASTER TABLE.csv")) %>% 
  select(`HS6 code`, `HS6 Description`, `FFD DESC`) %>% 
  group_by(`HS6 code`, `FFD DESC`) %>% 
  summarise(count = n()) %>% 
  filter(`FFD DESC` != "Not entered")

cncodes$`HS6 code`[duplicated(cncodes$`HS6 code`)]

```
